Scalar and Representative Observables, and Their Associated Statistical Model, for Polarimetric SAR Data

### 2014-08-10 paper_dev/article.tex
auto

\begin{abstract}
%Problem statement  
This paper proposes novel scalar and representative observables, including several statistically consistent discrimination measures,
  that would be useful in the interpretation and processing of multi-dimensional POLSAR data.
%What actually were done? and As the result, what is learned?
Specifically, the statistical behaviour of the POLSAR covariance matrix determinant is used
  to derive a scalar and generic statistical model for multi-dimensional POLSAR data,
  which is shown applicable to the two and three dimensional versions of partial and full monostatic polarimetric SAR data.
In addition, the proposed model is also able to subsume the traditional SAR intensity model under the umbrella of a unified model,
  where the SAR intensity is shown to be a one dimension version of the multi-dimensional data.
%As the POLSAR covariance matrix determinant generalizes the SAR intensity towards multiple dimensions,
%  the proposed model is able to subsume the traditional SAR intensity model under the umbrella of a unified model. 
%What are the larger implications of the findings presented in this paper?
Consequently, the main beneficial implications of the proposed approach are that
  it provides a consistent theory unifying the currently disconnected proposals for SAR and POLSAR discrimination measures,
  and it simplifies the adaptation of existing SAR data processing techniques for POLSAR data.
%  it enables the adaptation of many existing SAR data processing techniques for POLSAR data,
%  and at the same time, it provides a consistent theory unifying the seemingly disparate discrimination measure proposals for SAR and POLSAR data.
\end{abstract}

@@

Taking the log-transformation of the above equation, and denoting $C_{vx} = Z_x / L_x$, $C_{vy} = Z_y / L_y$ and $C_{vxy} = (Z_x + Z_y)/(L_x + L_y)$ then:
{\small
\begin{align}
  Q &= \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} % \\
  % \ln Q &= L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
\end{align}
}

@@

Note that the POLSAR scattering vector is denoted as $s$ in this paper.
For partial polarimetric SAR (single polarization in transmit and dual polarization in receive),
  the vector is two-dimensional ($d=2$) and is normally written as: 
\begin{equation}
s_{part}=\begin{bmatrix}
S_h\\ 
S_v
\end{bmatrix}
\end{equation}
For full and monostatic POLSAR data,
  the vector is three-dimensional ($d=3$) and is presented as:
\begin{equation}
s_{full}=\begin{bmatrix}
S_{hh}\\
\sqrt{2}S_{hv}\\
S_{vv}
\end{bmatrix}
\end{equation}

@@

The approach taken in this paper differs by applying the homoskedastic log transformation  on a less-than-well-known result for the determinant of the covariance matrix.
Goodman \cite{Goodman_1963_AMS_178} proved
that the ratio between the observable and expected values of the sample covariance matrix determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom: 
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

@@

%Assuming, on the one hand,
For cases where the true value of the underlying signal $\Sigma_v$ is known \textit{a priori},
then the determinant-ratio of the signal random variable ($\mathbb{R}_{\Sigma}$) %and log-distance ($\mathbb{L}$)
%  are observable according to the following definitions:
  is defined as:
%Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv} lead straight to the definition of the following random variables, which is the :
\begin{eqnarray}
  \mathbb{R}_{\Sigma} &=& \frac{|C_v|}{|\Sigma_v|} \label{eqn:determinant_ratio_observables}%\\
%  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
%On the other hand, under a looser assumption %From another perspective
For cases where the POLSAR data is known to have come from a homogeneous area, but the true value of the underlying signal $\Sigma_v$ is \textit{unknown},
 then a random variable called the change-ratio ($\mathbb{R}_{C}$)  is defined as:
\begin{eqnarray}
  \mathbb{R}_{C} &=& \frac{|C_1|}{|C_2|} \label{eqn:determinant_ratio_samples_observables}%\\
%  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
where $C_1$ and $C_2$ are samples of the covariance matrix determinant in an assumed homogeneous area.

@@

Upon setting $d=1$ into the above equations,
  the equations become:
\begin{align*}
  |C_v| &\sim  \frac{|\Sigma_v|}{(2L)} \cdot \chi (2L) \\ %\label{eqn:determinant_distribution} %\\
  \mathbb{R}_{\Sigma} = \frac{|C_v|}{|\Sigma_v|} &\sim \frac{1}{(2L)} \cdot \chi(2L) \\ %\label{eqn:polsar_ratio_det_cov_dist} \\
\mathbb{R}_{C} = \frac{C_1}{C_2} &\sim \frac{\chi(2L)}{\chi(2L)}
\end{align*}

Since the PDF of chi-squared distribution can be written as:
\begin{align*}
\chi(2L) \sim pdf \left[ \frac{\chi^{L-1}e^{-\chi/2}}{2^L\Gamma(L)} \right]
\end{align*}
Applying the variable change theorem into the above equations results in:
\begin{align*}
  |C_v| &\sim  pdf \left[ \frac{L^L x^{L-1} e^{-Lx/|\Sigma_v|}}{\Gamma(L) |\Sigma_v|^L} \right] \\ %\label{eqn:determinant_distribution} %\\
  \mathbb{R}_{\Sigma} = \frac{|C_v|}{|\Sigma_v|} &\sim pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \right] \\ %\label{eqn:polsar_ratio_det_cov_dist} \\
  \mathbb{R}_{C} = \frac{C_1}{C_2} &\sim pdf \left[ \frac{\Gamma(2L-1) x^{L-1}}{\Gamma^2(L-1) (1+x)^{2L}} \right]
\end{align*}
These equations match exactly with the following traditional model for multi-look SAR intensity:
  \begin{eqnarray}
I &\sim& pdf \left[ \frac{L^L x^{L-1} e^{-Lx/\bar{I}}}{\Gamma(L) \bar{I}^L} \right] \\
\mathbb{R}_{\bar{I}} = \frac{I}{\bar{I}} &\sim& pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \label{eqn:multi_look_SAR_ratio_dist} \right] \\
  \mathbb{R}_{I} = \frac{I_1}{I_2} &\sim& pdf \left[ \frac{\Gamma(2L-1) x^{L-1}}{\Gamma^2(L-1) (1+x)^{2L}} \right]
  \end{eqnarray}
considering that $|C_v| \mapsto I$ and $|\Sigma_v| \mapsto \bar{I}$ when multi-dimensional POLSAR collapses into single-dimensional SAR.

@@

As for the first matter, in view of the models given in Eqn \ref{eqn:determinant_distribution},
  the likelihood test statistics presented in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqns \ref{eqn:ori_likelyhood_stats} \& \ref{eqn:log_likelyhood_stats}
can be expressed as:
\begin{align*}
  %\ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
%where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
This, in essence, derives an exact statistical distribution for the likelihood test statistics (as opposed to an asymptotic distribution \cite{{Conradsen_2003_TGRS_4}}).

@@

Since it has been shown in Section \ref{sec:sar_special_case_of_polsar} that the model for the case of $d=1$ matches exactly with the traditional model for SAR intensity,
  the validation of the proposed $d=1$ model is quite straightforward. %only a simple validation is presented for this case.  
Fig. \ref{fig:verify_POLSAR_model_1D} presents the results of a test performed for the stated purpose.
In the test, the intensity of single-channel SAR data (HH) for sample homogeneous areas in both datasets is extracted.
The histograms for both the intensity and intensity ratio %the log-distance and contrast
  are then plotted against the theoretical PDF given above.
%Also presented are the plots of the same test applied on a homogeneous patch inside the nine-look processed RADARSAT2 image.
The plots are obtained having set ENL to an estimated value obtained using the technique presented in \cite{Anfinsen_2009_TGRS_3795}.
The values of $L$ are $L=3.2752$ and $L=3.4241$ for the AIRSAR and RadarSat2 images respectively.   
In all cases, the good match between the actual data and model distribution lends support to the proposed model. %providing us a simple validation of the result.

@@

The remainder of this section now focuses on validating the models for both partial ($d=2$) and full ($d=3$) POLSAR.
Similar tests %like the one described for SAR data 
  are carried out
  for both types of polarimetric SAR
  on both AIRSAR and RADARSAT2 datasets.
The look-number is estimated for each type of data and for each image~\cite{Anfinsen_2009_TGRS_3795}. %and is also noted in the plots.
The practical and predicted distributions are again plotted to determine whether their visual match is close in each case. Fig. \ref{fig:verify_det_ratio_model_2D} shows the $d=2$ plots for determinant, determinant ratio and change ratio for both datasets. It is very clear that the model histogram matches the data well, in fact appearing similar to a smoothed data response. Evidently the match is good for $d=2$.

@@

%conclude the section
Many similar SAR data processing techniques may be adaptable to POLSAR in this way, however they are not explored here due to space constraints.

@@

%What has been done?
This paper has proposed a generic statistical model  for the POLSAR covariance matrix determinant based on the complex Wishart POLSAR target vector distribution.
The model has been validated for specific $d=2$ and $d=3$ cases using real-life partial and full POLSAR data.
The paper also establishes two new POLSAR discrimination measures: the determinant-ratio and the change-ratio,  which essentially are the generic versions of the SAR intensity-ratio.
We have also shown an excellent match between the specific $d=1$ model and that for traditional SAR intensity,
which effectively brings the existing SAR theories under the umbrella of this new model.

>>>>

\begin{abstract}
%Problem statement  
This paper proposes novel scalar and representative observables, including several statistically consistent discrimination measures,
  that would be useful in the interpretation and processing of multi-dimensional POLSAR data.
%What actually were done? and As the result, what is learned?
Specifically, the statistical behaviour of the POLSAR covariance matrix determinant is used
  to derive a scalar and generic statistical model for multi-dimensional POLSAR data,
  which is shown applicable to the two and three dimensional versions of partial and full monostatic polarimetric SAR data.
In addition, the proposed model is also able to subsume the traditional SAR intensity model under the umbrella of a unified model,
  where the SAR intensity is shown to have similar behaviour to the one dimension version of the proposed multi-dimensional models.
%As the POLSAR covariance matrix determinant generalizes the SAR intensity towards multiple dimensions,
%  the proposed model is able to subsume the traditional SAR intensity model under the umbrella of a unified model. 
%What are the larger implications of the findings presented in this paper?
Consequently, the main beneficial implications of the proposed approach are that
  it provides a consistent theory unifying the currently disconnected proposals for SAR and POLSAR discrimination measures,
  and it simplifies the adaptation of existing SAR data processing techniques for POLSAR data.
%  it enables the adaptation of many existing SAR data processing techniques for POLSAR data,
%  and at the same time, it provides a consistent theory unifying the seemingly disparate discrimination measure proposals for SAR and POLSAR data.
\end{abstract}

@@

Taking the log-transformation of the above equation, and denoting $C_{vx} = Z_x / L_x$, $C_{vy} = Z_y / L_y$ and $C_{vxy} = (Z_x + Z_y)/(L_x + L_y)$ then:
{\small
\begin{align}
  Q &= \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} % \\
  % \ln Q &= L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
%  Q &= \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} \\
%  \ln Q &= L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
\end{align}
}

@@

Note that the POLSAR scattering vector is denoted as $s$ in this paper.
For partial polarimetric SAR (single polarization in transmit and dual polarization in receive), e.g. \cite{Souyris_2005_TGRS, Raney_2006_IGARSS}, 
  the vector is two-dimensional ($d=2$) and is normally written as: 
\begin{equation}
s_{part}=\begin{bmatrix}
S_h\\ 
S_v
\end{bmatrix}
\end{equation}
For full and monostatic POLSAR data,
  the vector is three-dimensional ($d=3$) and is presented as:
\begin{equation}
s_{full}=\begin{bmatrix}
S_{hh}\\
\sqrt{2}S_{hv}\\
S_{vv}
\end{bmatrix}
\end{equation}

@@

The approach taken in this paper differs 
  in that it make use of 
  %by applying the homoskedastic log transformation on 
  a less-than-well-known result for the determinant of the covariance matrix.
Goodman \cite{Goodman_1963_AMS_178} proved
that the ratio between the observable and expected values of the sample covariance matrix determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom: 
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

@@

%Assuming, on the one hand,
Assuming the true value of the underlying signal $\Sigma_v$ is \textit{available},
then the determinant-ratio of the signal random variable ($\mathbb{R}_{\Sigma}$) %and log-distance ($\mathbb{L}$)
%  are observable according to the following definitions:
  is defined as:
%Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv} lead straight to the definition of the following random variables, which is the :
\begin{eqnarray}
  \mathbb{R}_{\Sigma} &=& \frac{|C_v|}{|\Sigma_v|} \label{eqn:determinant_ratio_observables}%\\
%  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
%On the other hand, under a looser assumption %From another perspective
For cases where the POLSAR data is known to have come from a homogeneous area, but the true value of the underlying signal $\Sigma_v$ is \textit{unknown},
 then a random variable called the change-ratio ($\mathbb{R}_{C}$)  is defined as:
\begin{eqnarray}
  \mathbb{R}_{C} &=& \frac{|C_1|}{|C_2|} \label{eqn:determinant_ratio_samples_observables}%\\
%  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
where $C_1$ and $C_2$ are samples of the covariance matrix determinant in an assumed homogeneous area.

@@

Upon setting $d=1$ into the above equations,
  they become:
\begin{align*}
  |C_v| &\sim  \frac{|\Sigma_v|}{(2L)} \cdot \chi (2L) \\ %\label{eqn:determinant_distribution} %\\
  \mathbb{R}_{\Sigma} = \frac{|C_v|}{|\Sigma_v|} &\sim \frac{1}{(2L)} \cdot \chi(2L) \\ %\label{eqn:polsar_ratio_det_cov_dist} \\
\mathbb{R}_{C} = \frac{C_1}{C_2} &\sim \frac{\chi(2L)}{\chi(2L)}
\end{align*}

Since the PDF of chi-squared distribution can be written as:
\begin{align*}
\chi(2L) \sim pdf \left[ \frac{\chi^{L-1}e^{-\chi/2}}{2^L\Gamma(L)} \right]
\end{align*}
Applying the variable change theorem into the above equations results in:
\begin{align*}
  |C_v| &\sim  pdf \left[ \frac{L^L x^{L-1} e^{-Lx/|\Sigma_v|}}{\Gamma(L) |\Sigma_v|^L} \right] \\ %\label{eqn:determinant_distribution} %\\
  \mathbb{R}_{\Sigma} = \frac{|C_v|}{|\Sigma_v|} &\sim pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \right] \\ %\label{eqn:polsar_ratio_det_cov_dist} \\
  \mathbb{R}_{C} = \frac{C_1}{C_2} &\sim pdf \left[ \frac{\Gamma(2L-1) x^{L-1}}{\Gamma^2(L-1) (1+x)^{2L}} \right]
\end{align*}
These equations match exactly with the following traditional model for multi-look SAR intensity:
  \begin{eqnarray}
I &\sim& pdf \left[ \frac{L^L x^{L-1} e^{-Lx/\bar{I}}}{\Gamma(L) \bar{I}^L} \right] \\
\mathbb{R}_{\bar{I}} = \frac{I}{\bar{I}} &\sim& pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \label{eqn:multi_look_SAR_ratio_dist} \right] \\
  \mathbb{R}_{I} = \frac{I_1}{I_2} &\sim& pdf \left[ \frac{\Gamma(2L-1) x^{L-1}}{\Gamma^2(L-1) (1+x)^{2L}} \right]
  \end{eqnarray}
considering that $|C_v| \mapsto I$ and $|\Sigma_v| \mapsto \bar{I}$ when the proposed multi-dimensional POLSAR statistical models is applied towards the case of single-dimensional SAR.

@@

As for the first matter, in view of the models given in Eqn \ref{eqn:determinant_distribution},
  the likelihood test statistics presented in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqn. \ref{eqn:ori_likelyhood_stats} %\& \ref{eqn:log_likelyhood_stats}
can be expressed as:
\begin{align*}
%  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
%where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
This, in essence, derives an exact statistical distribution for the likelihood test statistics (as opposed to an asymptotic distribution \cite{{Conradsen_2003_TGRS_4}}).

@@

Since it has been shown in Section \ref{sec:sar_special_case_of_polsar} that the model for the case of $d=1$ matches exactly with the traditional model for SAR intensity,
  the validation of the proposed $d=1$ model is quite straightforward. %only a simple validation is presented for this case.  
Fig. \ref{fig:verify_POLSAR_model_1D} presents the results of a test performed for the stated purpose.
In the test, the intensity of single-channel SAR data (HH) for sample homogeneous areas in both datasets is extracted.
The sizes are 300x300 and 50x50 for Radarsat2 and AIRSAR patches respectively. 
The histograms for both the intensity and intensity ratio %the log-distance and contrast
  are then plotted against the theoretical PDF given above.
%Also presented are the plots of the same test applied on a homogeneous patch inside the nine-look processed RADARSAT2 image.
The plots are obtained having set ENL to an estimated value obtained using the technique presented in \cite{Anfinsen_2009_TGRS_3795}.
The values of $L$ are $L=3.2752$ and $L=3.4241$ for the AIRSAR and RadarSat2 images respectively.   
In all cases, the good match between the actual data and model distribution lends support to the proposed model. %providing us a simple validation of the result.

@@

The remainder of this section now focuses on validating the models for both partial ($d=2$) and full ($d=3$) POLSAR.
Similar tests %like the one described for SAR data 
  are carried out
  for both types of polarimetric SAR
  on both AIRSAR and RADARSAT2 datasets.
The look-number is estimated for each type of data and for each image~\cite{Anfinsen_2009_TGRS_3795}. %and is also noted in the plots.
The practical and predicted distributions are again plotted to determine whether their visual match is close in each case. Fig. \ref{fig:verify_det_ratio_model_2D} shows the $d=2$ plots for determinant, determinant ratio and change ratio for both datasets. 
It is clear that the model histogram matches the data well
  %, in fact appearing similar to a smoothed data response. 
Evidently the match is good for $d=2$.

@@

%conclude the section
Many similar SAR data processing techniques may be adaptable to POLSAR in this way, however they are not explored exhaustively here due to space constraints.

@@

%What has been done?
This paper has proposed a generic statistical model  for the POLSAR covariance matrix determinant based on the complex Wishart POLSAR target vector distribution.
The model has been validated for specific $d=2$ and $d=3$ cases using real-life partial and full POLSAR data.
The paper also establishes two new POLSAR discrimination measures: the determinant-ratio and the change-ratio,  which essentially are the generic versions of the SAR intensity-ratio.
We have also shown an excellent match between the specific $d=1$ model and that of the traditional SAR intensity,
which effectively brings the existing SAR theories under the umbrella of this new model.

### 2014-08-10 paper_dev/article_pp.tex
auto

As for the first matter, in view of the models given in Eqn \ref{eqn:determinant_distribution},
  the likelihood test statistics presented in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqns \ref{eqn:ori_likelyhood_stats} \& \ref{eqn:log_likelyhood_stats}
can be expressed as:
\begin{align*}
%  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
%where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
This, in essence, derives an exact statistical distribution for the likelihood test statistics (as opposed to an asymptotic distribution \cite{{Conradsen_2003_TGRS_4}}).

>>>>

As for the first matter, in view of the models given in Eqn \ref{eqn:determinant_distribution},
  the likelihood test statistics presented in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqn. \ref{eqn:ori_likelyhood_stats} %\& \ref{eqn:log_likelyhood_stats}
can be expressed as:
\begin{align*}
%  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
%where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
This, in essence, derives an exact statistical distribution for the likelihood test statistics (as opposed to an asymptotic distribution \cite{{Conradsen_2003_TGRS_4}}).
